---
title: "Advanced Graphs"
author: "Barry DeCicco"
date: "2025-09-11"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Contents


This presentation covers various useful plot tips I've picked up from various sources.   It's intended to make your life a bit simpler. My sources are the R Graph Gallery, LinkedIn posts and the Medium website.

## The SanKey Diagram

The SanKey diagram is used to illustrate the low of quantities from earlier sources to later (or intermediate) destinations.

Sankey diagrams are named after Irish Captain Matthew Henry Phineas Riall Sankey, who used this type of diagram in 1898 in a classic figure showing the energy efficiency of a steam engine. The original charts in black and white displayed just one type of flow (e.g. steam); using colors for different types of flows lets the diagram express additional variables. (See: https://en.wikipedia.org/wiki/Sankey_diagram#cite_note-3)

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## References:

- The R Graph Gallery (https://r-graph-gallery.com)
- Wikipedia (https://en.wikipedia.org/wiki/Main_Page)
